Superphy: Real-time prediction of \textit{E. coli} phenotype using graph-based genome sequence storage and retrieval with the \textit{spfy} module

### 2017-11-13 paper-webserver.tex
EDIT: merged bullets into the into paras

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} virulence gene annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}. Comprehensive platforms that combine individual programs into a cohesive whole also exist. These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC. Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests {REF}.

@@

% 1. Our specific problem, previous work
We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}. Superphy integrates pre-computed results with domain-specific knowledge to provide real-time exploration of publicly available genomes. While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.

% 2. Why solving it is important (for Public health / research)
WGS is becoming the norm for bacterial surveillance, outbreak tracing, and research \cite{ronholm2016navigating}. While pipelines exist \cite{joensen2014real} for running a variety of analyses on WGS data, 
In order to  Existing annotation and analysis software for WGS data represent units of compuatation; we can optimize WGS pipelines if these results are stored to avoid recomputation, and extend results by viewing them in the context of other analysis modules and genomes.
 these computations can be 
- everything is being sequenced (surveillance / outbreak / research)
- previously mentioned common analyses / want to leverage pre-computed results
- WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
- need fast / standardized outbreak response
- need fast / standardized in-silico reference lab
- need fast / standardized storage and retrieval of results based on ontology
- all known data can be leveraged, allows the most informed decisions possible
- etc.

% 3. how we solved it
In this study, we present an update to the SuperPhy platform, called Spfy. Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph data structures for further comparative genomics. Unlike traditional approaches which store results in separate data files or in a relational database, the graph data structures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future. We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.

>>>>

% 2. Big Problem: but tools are for individual analysis
The rapid characterization and comparison of bacterial pathogens relies principally on the combination of outputs from multiple software programs that are targeted for specific applications. Examples include the Resistance Gene Identifier (RGI) \cite{mcarthur2013comprehensive} for antimicrobial resistance (AMR) gene prediction, VirulenceFinder \cite{kleinheinz2014applying} virulence gene annotation, and Prokka for bacterial genome annotation with external tools \cite{doi:10.1093/bioinformatics/btu153}. In particular, RGI and VirulenceFinder represent a series of \textit{in-silico} methods which have been developed to mimick the results of traditional tests. These methods allow new WGS results to be viewed in the context of historical tests.

Comprehensive platforms that combine individual programs into a cohesive whole also exist. These include free platforms such as the Bacterium Analysis Pipeline (BAP) \cite{thomsen2016bacterial}, the Integrated Rapid Infectious Disease Analysis (IRIDA) project \url{http://www.irida.ca/}, and PATRIC. Commercial applications, such as Bionumerics, which is used by PulseNet international for the analyses of WGS data in outbreak situations also exist, and offer support as well as accredited, standardized tests {REF}.

@@

% 4. Our specific problem, previous work
We have previously developed Superphy \cite{whiteside2016superphy}, an online predictive genomics platform targeting \textit{E. coli}. Superphy integrates pre-computed results with domain-specific knowledge to provide real-time exploration of publicly available genomes. While this tool has been useful for the thousands of pre-computed genomes in its database, the current pace of genome sequencing requires real-time predictive genomic analyses of tens-, and soon hundreds-of-thousands of genomes, and the long term storage and referencing of these results, something that the original SuperPhy platform was incapable of.

% ?. Why solving it is important (for Public health / research)
% Merging points from here into the above sections (2,3,4) to avoid repetition
% - everything is being sequenced (surveillance / outbreak / research)
% - previously mentioned common analyses / want to leverage pre-computed results
% - WGS does not discard old methods, linkage to thousands of historical results by developing in-silico methods of traditional tests
% - need fast / standardized outbreak response
% - need fast / standardized in-silico reference lab
% - need fast / standardized storage and retrieval of results based on ontology
% - all known data can be leveraged, allows the most informed decisions possible
% - etc.

% 5. how we solved it
In this study, we present an update to the SuperPhy platform, called Spfy. Spfy enables large-scale genome processing in real-time and uniquely decomposses results into graph data structures for further comparative genomics. Unlike traditional approaches which store results in separate data files or in a relational database, the graph data structures are then integrated into a knowledge graph identifying relationships between all genomes sequenced in the past, present, and future. We present the use of graph databases as a solution to storing genome analysis results in a meaningful manner, and to allow retrospective comparisons as more genomes are sequenced or populations change.
